---
title: "S&DS 230 Project"
author: "Matt Shu, Cem Kupeli, Shen-Vey Lai, Se Ri Lee"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

*Music is a big part of our lives! We stream music all the time, and we occasionally peek at charts to see if our favorite artist made the list. We may also wonder how certain variables, like rank, total number of listeners, and length of lyrics, may be related to each other. For our report, we specifically looked at Last.fm. We wanted to see if the total number of listeners for each song can be predicted by variables such as rank, length of lyrics, year, and the number of artist appearances.*

## Data

* Response Variable
    + Number of listeners on Last.fm (i.e. the cumulative number of listeners for each track since the inception of Last.fm or the release of the track)
* Categorical Variables
    + Gender 
    + Region (i.e. the country where the song was produced)
    + Type (band or soloist)
* Continuous Variables 
    + Year (i.e. the year in which the chart rankings are taken from)
    + Rank (i.e. a value between 1-100 representing its place the first time it appears on a year-end chart)
    + forcast (i.e. a measure of a lyric’s readability measured by the proportion of one syllable words in a song)
    + mean_word_syllables (i.e. the average number of syllables per word in the lyrics) 

## Prereqs

```{r, message=FALSE, results = 'hide', warning=FALSE}
library(tidyverse)
library(RSQLite)
library(RecordLinkage)
library(stringdist)
library(devtools)
library(quanteda)
library(quanteda.textstats)
library(car)
library(leaps)
library(corrplot)
library(PerformanceAnalytics)
source("http://www.reuningscherer.net/s&ds230/Rfuncs/regJDRS.txt")
```

```{r}
conn <- dbConnect(RSQLite::SQLite(), "22-04-29-playback-fm-top-pop.db")
dfSongs <- dbGetQuery(conn, 'SELECT * FROM tracks')
dim(dfSongs)
dfArtists <- dbGetQuery(conn, 'SELECT * FROM artists')
dbDisconnect(conn)
```

*We used RSQLite to separate the raw data into two dataframes: dfSongs and dfArtists. dfSongs contains information about the songs, and dfArtists stores information about the artists.*

## Data Cleaning
### Artist Dataset
#### Cleaning + Subset of Interest
```{r}
dfArtists[dfArtists == "nan"] <- NA
dfArtistsInterest <- dfArtists %>%
   dplyr::select(artist_id, type, area.name, gender) %>%
   as.data.frame()
```

*Then, we created a new dataframe dfArtistsInterest, which is a subset of the dataframe dfArtists. dfArtistsInterest contains four columns: the id of the artist, the type (whether the artist is a band or soloist), the area/region the song was produced/released, and the gender of the artist. *

#### Merge Artists with Songs
```{r}
dfSongsArtists <- merge(dfSongs,dfArtistsInterest,by="artist_id")
dim(dfSongsArtists)
```

*We then merged the dfSongs dataframe with the dfArtistsInterest dataframe.*

### Filter out NA's in Important Parameters
```{r Remove Missing}
cleaned_df <-dfSongsArtists %>%
  filter(!is.na(lyrics))  %>%
  filter(!is.na(artist))  %>%
  filter(!is.na(area.name)) %>%
  filter(!is.na(type))  %>%
  filter(!is.na(last_fm_listeners))  %>%
  as.data.frame()
dim(cleaned_df)
```

*In cleaned_df, we kept only the rows in which lyrics, artist, area.name, type, and last_fm_listeners are not NA.*

### Gender
```{r, results='hide'}
unique(cleaned_df$gender)
cleaned_df <- cleaned_df %>% 
    mutate(gender = replace(gender, gender == "other", "non-binary"))
cleaned_df <- cleaned_df %>% 
    mutate(gender = replace(gender, is.na(gender), "group"))
cleaned_df %>% count(gender)
cleaned_df <- cleaned_df %>% filter(gender != "non-binary")
dim(cleaned_df)
```

*In this section, we cleaned the variable Gender. Before cleaning, the possible values for Gender are NA, male, female, other, and non-binary. We changed all “other” values to “non-binary.” Then, for the NAs, we changed their value to “group” since a gender of NA probably means the artist is a band, not a soloist. We found that most artists were bands (groups). Out of the soloists, male artists were the most prevalent. Since the number of non-binary singers is small, we unfortunately removed them from our dataset.*

### area.name -> country
```{r}
cleaned_df <- cleaned_df %>% 
    mutate(country = dplyr::recode(area.name, "Los Angeles" = "United States", "Boston" = "United States", "Malvern" = "United States", "Olympia Fields" = "United States", "Alpharetta" = "United States", "Atlanta" = "United States", "Nordrhein-Westfalen" = "Germany", "Saddle River" = "United States", "Florida" = "United States", "Hollywood" = "United States", "Manhattan" = "United States", "Vancouver" = "Canada", "Portland" = "United States", "Quebec" = "Canada", "New York" = "United States", "Hawaii" = "United States", "Devon" = "United States", "Toronto" = "Canada", "[Worldwide]" = "Worldwide", "London" = "United Kingdom", "British Virgin Islands" = "United Kingdom", "Brooklyn" = "United States", "Ann Arbor" = "United States", "Salt Lake City" = "United States", "Rome" = "Italy", "Nashville" = "United States", "Chicago" = "United States", "Houston" = "United States", "Scotland" = "United Kingdom", "England" = "United Kingdom", "Puerto Rico" = "United States"))
cleaned_df <- cleaned_df %>% 
    mutate(region = dplyr::recode(country, "Austria" = "Non-UK Europe", "Belgium" = "Non-UK Europe", "Austria" = "Non-UK Europe", "Denmark" = "Non-UK Europe", "Finland" = "Non-UK Europe", "France" = "Non-UK Europe", "Germany" = "Non-UK Europe", "Greece" = "Non-UK Europe", "Iceland" = "Non-UK Europe", "Ireland" = "Non-UK Europe", "Italy" = "Non-UK Europe", "Moldova" = "Non-UK Europe", "Netherelands" = "Non-UK Europe", "Norway" = "Non-UK Europe", "Romania" = "Non-UK Europe", "Russia" = "Non-UK Europe", "Switzerland" = "Non-UK Europe", "Sweden" = "Non-UK Europe", "Spain" = "Non-UK Europe", "Netherlands" = "Non-UK Europe", "Senegal" = "Africa", "Guinea" = "Africa", "Jamaica" = "Africa", "Morocco" = "Africa", "South Africa" = "Africa", "Bahamas" = "Non-Canada/US Americas", "Barbados" = "Non-Canada/US Americas", "Jamaica" = "Non-Canada/US Americas", "Colombia" = "Non-Canada/US Americas", "Panama" = "Non-Canada/US Americas", "Saint Vincent and The Grenadines" = "Non-Canada/US Americas", "Japan" = "Misc", "South Korea" = "Misc", "Worldwide" = "Misc", "Philippines" = "Misc", "Australia" = "Oceania", "New Zealand" = "Oceania"))
# cleaned_df %>% count(region)
cleaned_df <- cleaned_df %>% 
    mutate(region = dplyr::recode(region, "Africa" = "Misc", "Non-Canada/US Americas" = "Misc", "Oceania" = "Misc"))
cleaned_df %>% count(region)
```

*Next, we condensed the variable Region so as to limit the number of different categories. We started by replacing the names of cities and states in the variable to their respective countries (e.g: Boston → United States). We then went on to categorize countries with less frequent appearances in the data into their larger affiliated regions (e.g: Austria → Non-UK Europe) and others without clear regional affiliations into Misc. Eventually, we decided to combine the regions Africa, Non-Canada/US Americas and Oceania and place them under Misc as well. By the end of this, we ended up with 5 main regions: Canada, United States, United Kingdom, Non-UK Europe and Misc.*

### Type
```{r}
unique(cleaned_df$type)
cleaned_df %>% count(type)
dim(cleaned_df)
cleaned_df <- cleaned_df %>% filter(type %in% c("Group", "Person"))
dim(cleaned_df)
```

*For this step, we wanted to categorize the types of artists into two distinct types: soloists (Person) and bands (Group). However, while doing this, we encountered a bug in which, despite cleaning the data so that the variable Type would only contain Person or Group, there was one row that had “Other” as its data Type. We took out this "Other," though we were unable to figure out why it was there.*

### Filter out Mismatches in Lyrics

```{r Remove Erroneous Lyrics}
# 
cleaned_df$lyrics <- str_replace_all(cleaned_df$lyrics,"[\\s]+", " ")
cleaned_df$cleaned_lyrics <- 
  str_replace_all(cleaned_df$lyrics, 'Chap\\. [0-9]', NA_character_) %>%
  str_replace_all(., 'Listening Log', NA_character_) %>%
  str_replace_all(., 'Favorite Songs Of', NA_character_) %>%
  str_replace_all(., 'Chapter [0-9]', NA_character_) %>%
  str_replace_all(., 'New Music ', NA_character_) %>%
  str_replace_all(., 'Nominees', NA_character_) %>%
  str_replace_all(., 'Best Songs of ', NA_character_) %>%
  str_replace_all(., "[0-9]+ U S", NA_character_) %>% # Court Cases
  str_replace_all(., "[0-9]+ U.S", NA_character_) %>% # Court Cases
  str_replace_all(.,"[ ]+", " ") %>%
  str_replace(., ".*Lyrics", "") %>%
  str_replace(., "[0-9]*Embed$", "")

cleaned_df <- cleaned_df %>%
  filter(!is.na(cleaned_lyrics)) %>%
  filter(levenshteinSim(track, str_match(lyrics, "(.*)Lyrics")[,2]) > .5) %>% # There are some false positives, when there are other languages
  as.data.frame()

# Filter away songs with less than 100 chars, must have been a mistake in lyrics
cleaned_df <- cleaned_df %>% filter(nchar(cleaned_lyrics) > 100)
dim(cleaned_df)
```

*Now on to cleaning the lyrics! For this, we started by identifying the “problematic” lyrics in the data, particularly ones that did not contain the actual lyrics. We then replaced these strings with NA characters and proceeded to remove them. Another method we used to clean the lyrics was to use the levenshteinSim function. The variable lyrics is formatted such that it should start with the name of the track, followed by the string ‘Lyrics’ and then the actual lyrics of the song. The function allowed us to compare for similarity between the title of the track and the initial part of the “lyrics”. If the two strings were not at least 50% similar, we removed them.*

### Readability
```{r}
cleaned_corpus <- corpus(
  cleaned_df,
  docid_field = "track_id",
  text_field = "cleaned_lyrics",
  unique_docnames = TRUE
)
```

*Then, we quantified the variable lyrics so that we could use it for our data analysis. One such method is to use the function textstat_readability, which could only read in the corpus data type. Thus, we used the corpus function to convert our cleaned_df dataframe into a corpus data type.*

```{r, results = F}
# We use FORCAST because it's one of the only readability metrics that don't look for sentences
cleaned_df$forcast <- textstat_readability(
  cleaned_corpus,
  measure = "FORCAST",
  min_sentence_length = 3,
  max_sentence_length = 10000,
)$FORCAST
cleaned_df$mean_word_syllables <- textstat_readability(
  cleaned_corpus,
  measure = "meanWordSyllables",
  remove_hyphens = TRUE,
  min_sentence_length = 3,
  max_sentence_length = 10000,
)$meanWordSyllables
```

*Then, using the function textstat_readability, we calculated the FORCAST readability of the lyrics, which is basically the proportion of one syllable words to the total number of words. We chose the FORCAST metric over other metrics because it is one of the only metrics that don’t look for sentences. Since most lyrics do not contain fully formed grammatical sentences with periods, the FORCAST metric is suitable for our data. We additionally add a second readability metric, which is simply the mean number of syllables for a word in a song.*

### Filter to Relevant Details

```{r}
subset_df <- cleaned_df %>%
   dplyr::select(year, type, region, gender, last_fm_listeners, artist_appearances, rank, forcast, mean_word_syllables)
subset_df$log_listeners <- log(subset_df$last_fm_listeners)
attach(subset_df)
```

## Descriptive Plots
### Histograms
```{r}
hist(last_fm_listeners, main="Histogram of Last.fm listeners", xlab="Last.fm listeners", ylab = "Frequency", cex.main = 1, cex.lab = 1, col = "blue")
```

*The histogram of last.fm listeners is heavily right-skewed.*

```{r}
hist(log_listeners, main="Histogram of Log Last.fm listeners", xlab="Last.fm listeners", ylab = "Frequency", cex.main = 1, cex.lab = 1, col = "blue")
```

*Thus, we applied a log transformation on last.fm listeners to see if it might solve the skew problem. However, the log of last.fm listeners turns out to be left-skewed. This suggests we might need to do a different transformation on last.fm listeners.*

### Scatterplots
```{r}
plot(last_fm_listeners ~ rank, main = "Last.fm listeners versus Rank", xlab = "Rank", ylab = "Last.fm listeners", cex.main = 1, 
     cex.lab = 1, col = "red")
```

*In this scatterplot, it appears that most of the data seems to be concentrated around the less than 500,000 number of listeners mark. There also does not appear to be any indication of a clearly defined relationship between last.fm listeners and rank. However, when looked at carefully, there seems to be a slight downward trend as the rank increases in numeric value (i.e. as the song places lower on the chart). It is also quite interesting that there are data points beyond the 1 million mark for ranks numerically above 50, which is surprising considering how we might assume that songs lower on the chart ranks would have fewer cumulative listeners. This might hint to what we could refer to as the “staying power” of a song, which refers to the popularity of the song across its existence on last.fm.*

### Boxplots
```{r}
boxplot(mean_word_syllables ~ gender, main = "Boxplot of Mean word syllables by Gender", ylab = "Mean word syllables", xlab = "Gender", cex.main = 1,
        cex.lab = 1, col = "yellow")
```
*The boxplots of mean word syllables by gender show that the mean word syllables across genders have similar medians and interquartile ranges. The genders Group (Band) and Male, however, seem to have a lot more outliers than Female.*

## Analysis/Tests

### Normal quantile plots
```{r}
qqPlot(log_listeners, xlab = "Normal Quantiles", ylab = "Log of Last.fm listeners", 
       main = "Normal Quantile Plot of Log of Last.fm listeners")
```
```{r}
qqPlot(last_fm_listeners, xlab = "Normal Quantiles", ylab = "Number of Last.fm listeners", 
       main = "Normal Quantile Plot of Number of Last.fm listeners")
```

### Correlation between number of listeners and rank (correlation, bootstrapped correlation)

```{r}
# Calculate and report the correlation along with the results of a parametric test of the significance of the correlation
(cor1 <- cor(last_fm_listeners, rank))
cor1test <- cor.test(last_fm_listeners, rank)

# Calculate a 95% bootstrap confidence interval for the true correlation
N <- 10000
cors <- rep(NA, N)
for (i in 1:N)
{
  # listeners and rank have the same length, so we can use either
  s <- sample(1:length(last_fm_listeners), length(last_fm_listeners), replace = TRUE)
  
  cors[i] = cor(last_fm_listeners[s], rank[s])
}
corsCI <- quantile(cors, c(0.025, 0.975))
print("95% confidence interval for the true correlation")
corsCI

# Display the results on a histogram (bootstrapped sample correlations, bootstrap confidence interval, and theoretical confidence interval)
hist(cors, main = "Bootstrap Correlations between Number of listeners and Rank", col = "blue")
abline(v = corsCI, lwd = 3, col = "red")
abline(v = cor1test$conf.int, lwd = 3, col = "green", lt = 2)
legend("topright", c("Theoretical CI", "Bootstrap CI"), lwd = 3, col = c("green", "red"), lty = c(2, 1))
```

*There is a weak negative correlation between the raw number of listeners and rank. We also calculated the bootstrapped correlation between these two variables. The bootstrapped confidence intervals and theoretical confidence intervals appear to be the same, though the theoretical CI is slightly narrower.*

### Difference in FORCAST readibility between male and female artists (t-test)
```{r}
# Check the normality of FORCAST data
qqPlot(forcast, main = "Normal Quantile Plot of FORCAST readibility scores", ylab = "FORCAST readibility score", xlab = "Normal Quantiles")

# The normal quantile plot is not exactly linear, so we perform a Box Cox transformation
trans1 <- boxCox(aov(forcast[gender != "group"] ~ gender[gender != "group"]))
pow1 <- trans1$x[which.max(trans1$y)]

# Apply the transformation to FORCAST
trans_forcast <- forcast^-1

# Omit "Group" as a type to compare female and male artists
test1 <- t.test(forcast[gender != "group"] ~ gender[gender != "group"])
test1
```

*We then wanted to see if there was a significant difference in FORCAST readability between male and female artist produced songs. First, we checked to see if FORCAST is normally distributed. The normal quantile plot is not linear, suggesting that FORCAST is not normally distributed. Thus, we performed a box-cox transformation and got a lambda of around -1. Thus, we transformed the FORCAST variable by raising it to -1 power. The p-value from the Welch’s t-test is 0.011, which is less than 0.05. Thus, we can reject the null hypothesis and conclude that there is a significant difference in the transformed FORCAST readability between male and female artists.*

### Difference in log of number of listeners between female and male artists (permutation test)

```{r}
# Transformed version of gender containing only solo artists (omitting the type "group")
soloGen <- gender[gender != "group"]

actualDiff <- median(log_listeners[soloGen == "female"]) - median(log_listeners[soloGen == "male"])

diffValues <- rep(NA, N)
for (i in 1:N)
{
  fakeGender <- sample(soloGen)
  diffValues[i] <- median(log_listeners[fakeGender == "female"]) - median(log_listeners[fakeGender == "male"])
}

hist(diffValues, xlab = "Log of number of listeners", main = "Difference in median log of number of listeners between females and males", col = "red",
     breaks = 20, cex.main = 0.7)
abline(v = actualDiff, col = "blue", lwd = 3)
text(actualDiff + 0.015, 1000, paste("Actual difference: ", round(actualDiff, 3)), srt = 90, cex = 0.9)

mean(abs(diffValues) >= abs(actualDiff))
```

*The p-value is greater than any alpha value, so we fail to reject the null hypothesis. Thus, there is no significant difference in the log number of listeners between male and female artists.*

### Mixed Correlation Plots
```{r}
# Calculate the pairwise correlations
cor1 <- cor(subset_df[ , c("log_listeners", "year", "artist_appearances", "rank", "forcast", "mean_word_syllables")])

# Calculate the pairwise correlation significances
sigcorr <- cor.mtest(subset_df[ , c("log_listeners", "year", "rank", "artist_appearances", "forcast", "mean_word_syllables")], conf.level = .95)

#Use corrplot.mixed to display confidence ellipses, pairwise correlation values, and put on 'X' over non-significant values.
corrplot.mixed(cor1, lower.col="black", upper = "ellipse", tl.col = "black", number.cex=.7, 
                tl.pos = "lt", tl.cex=.7, p.mat = sigcorr$p, sig.level = .05)

chart.Correlation(subset_df[ , c("log_listeners", "year", "rank", "artist_appearances", "forcast", "mean_word_syllables")], histogram = TRUE, pch = 19)
```

*The only somewhat strong(er) positive correlation that exists is between the log number of listeners and the year. On the other hand, the only somewhat strong(er) negative correlation is between the log number of listeners and the rank. There also appears to be a highly strong correlation between forcast and mean word syllables, which makes sense since the two predictor variables are directly related. This is an example of collinearity.*

### Box Cox transformation
```{r}
# Perform a Box Cox transformation
trans2 <- boxCox(lm(last_fm_listeners ~ gender + year + rank + region + artist_appearances + gender*year + year*rank + region*gender, data = subset_df))
pow2 <- trans2$x[which.max(trans1$y)]

# Apply the transformation to the variable
trans_listeners <- last_fm_listeners^pow2

hist(trans_listeners, main="Histogram of Last.fm listeners (transformed)", xlab="Last.fm listeners ^ 0.303", ylab = "Frequency", cex.main = 1, cex.lab = 1, col = "blue")
```

*Earlier, we found that both the raw data and log transformed data of last.fm listeners produced skewed histograms. Here, we performed a boxcox transformation on last.fm listeners to hopefully obtain a normal distribution. We got a lambda value of 0.303, and so we raised last.fm listeners to the 0.303 power. The histogram now looks a lot more normally distributed.*

### ANCOVA
```{r}
anc1 <- lm(trans_listeners ~ year + type + year*type)
Anova(anc1, type = 3)
summary(anc1)
myResPlots2(anc1)
```

*We then performed ANCOVA to predict the transformed last.fm listener variable with the variables year, type, and the interaction between year and type. Based on their p-values which are all lower than 0.05, we can conclude that year, type, and the interaction between year and type are all significant predictors. The model fit is great. The normal quantile plot of the residuals is linear, suggesting that the residuals are normally distributed. The fits vs residuals plot shows no heteroskedasticity and contain minimal outliers.*

*Lastly, we performed a backwards stepwise regression, a kind of multiple regression, to assess which variables and their interactions are significant predictors of the transformed last.fm listeners. We included the variables gender, year, rank, region, and the number of artist appearances, as well as the interaction between gender and year, year and rank, and region and gender. Due to the strong positive correlation between mean_word_syllables and FORCAST, we omitted mean_word_syllables before performing backwards stepwise regression. We used a significance of 0.001 since most of the p-values are very low to begin with. We started out by looking at the p-values of the interaction terms, eliminating first the interaction term with the highest p-value.*

### GLM and Backwards Stepwise Regression
```{r}
mod1 <- lm(trans_listeners ~ gender + year + rank + region + artist_appearances + gender*year + year*rank + region*gender)
Anova(mod1, type = 3)

mod2 <- lm(trans_listeners ~ gender + year + rank + region + artist_appearances + gender*year + region*gender)
summary(mod2)
myResPlots2(mod2)
```

*The interaction between year and rank had a p-value of 0.01, greater than our significance 0.001. Thus, we took this term out of our model. In our new model, the remaining two interaction terms had p-values less than 0.001. Thus, we kept these two terms and exitted out of the backwards regression. Our final model thus contains all variables we started out with except for the interaction between year and rank. The R-squared of our final model is 0.2502. Thus, about 25% of the variability in transformed last.fm listeners can be explained by the predictor variables. The normal quantile plot of the studentized residuals is linear, suggesting that the residuals are normally distributed. The fits vs studentized residuals shows no heteroskedasticity and contain minimal outliers. Thus, our final model fit is great!*

## Conclusion

*Overall, we did not find any strong relationship between any of the variables, other than forcast and mean_word_syllables which are obviously related. However, there was a weak negative correlation between the number of unique listeners on last.fm and the rank of the tracks. That is, the lower the rank of the song (or the higher the numeric value of the ranks), the lower the number of unique listeners, which makes intuitive sense. It is also interesting that the FORCAST readability between male and female artists is significantly different. Although we cannot make any strong statements based on these correlations, our results suggest that there are many other variables at play that we have not considered, many of which are probably not available to the public. Determining which tracks make the chart and which don't is a tricky and complex process that must take into account a multitude of factors.*






